Museum specimens as baseline for assessing a potential global pollinator crisis

### 2018-03-08 ms.Rmd
updated ms

Human colonisation of New Zealand was relatively recent in global terms (c. 740 y) (@Wilmshurst et al. 2008). Before human arrival, New Zealand was mostly forested, but has since been dramatically altered by people. Early Māori settlers cleared forests by burning and more recently, European colonists cleared much of remaining forests for agriculture, mostly before 1900 (@Perry et al, 2014). Therefore, human activity likely affected New Zealand pollinator communities long before collections began. Nevertheless, we can use museum records to identify trends in pollinator communities during New Zealand’s more recent history.

@@

Data analyses
First, we filtered our original datasets so that data used for analyses only included unique collection events. To do this, we removed specimens collected at the same location, on the same data and by the same collector. We found our data had reasonable coverage across time periods, although there was a peak in collection occurrences from 1960-1980. We accounted for differences in collection effort through time by binning collection records, so that each bin had a similar number of records but a different number of years (see @Bartomeusetal2013 for further details). We then estimated richness for each time period bin by rarefying all bins to an equal number of specimens and calculated the mean species richness ±SE for each bin. Finally, we estimated the significance of change in richness using a permutation test that randomly reordered time periods and calculated the correlation between time period and species richness. Thus reported P-values were the proportion of permutations that had higher or lower correlations compared to the correlation between richness and the actual chronological time period sequence.

To test whether there were relative differences in occurrences of species over time (should we analyse natives and exotics together or separately?). To determine if the probability of finding a species in the collection changed over time we used a general linear model with a binomial distribution and a logit link. For species that showed overdispersion, we used a quasi-binomial distribution. We only included species in this analysis that had more than 30 records. To account for differences in sampling effort between years, we weighted each year by the total number of samples collected in that year.

>>>>

Human colonisation of New Zealand was relatively recent in global terms (c. 740 y) (@Wilmshurst et al. 2008). Before human arrival, New Zealand was mostly forested, but has since been dramatically altered by people. Early Māori settlers cleared forests by burning and more recently, European colonists cleared large tracts of remaining forests and drained low-lying wetlands for agriculture, mostly before 1900 (@Perry et al, 2014). Therefore, human activity likely affected New Zealand pollinator communities long before surveys and collections of pollinators began. Nevertheless, we can use museum records to identify trends in pollinator communities during New Zealand’s more recent history.

@@

Data analyses
First, we filtered our original datasets so that data for analyses only included unique collection events. To do this, we removed specimens collected at the same location, on the same date, and by the same collector. We found our data had reasonable coverage across time periods, although there was a peak in collection occurrences from 1960-1980. We accounted for differences in collection effort through time by binning collection records, so that each bin had a similar number of records but a different number of years (see @Bartomeusetal2013 for further details). We then estimated richness for each time period bin by rarefying all bins to an equal number of specimens and calculated the mean species richness ±SE for each bin. Finally, we estimated the significance of change in richness using a permutation test that randomly reordered time periods and calculated the correlation between time period and species richness. Thus reported P-values were the proportion of permutations that had higher or lower correlations compared to the correlation between richness and the actual chronological time period sequence.

To determine if the probability of finding a species in the collection changed over time, we used a general linear model with a binomial distribution and a logit link. (should we analyse natives and exotics together or separately?). For species that showed overdispersion, we used a quasi-binomial distribution. We only included species in this analysis that had more than 30 records. To account for differences in sampling effort between years, we weighted each year by the total number of samples collected in that year.
